On the spatial aggregation of ecosystem condition indicators

### 2024-11-20 manuscript/manuscript.qmd
add normalise early

### Alien species example

@@

### Phytoplankton trophic index (PTI) {#sec-pti}

@@

### 1. Always report the aggregation pathway, using standardised terminology {.unnumbered}

@@

### 2. Always normalise variables at the scale where the reference levels are relevant {.unnumbered}

Normalising variables early in the aggregation pathway has several benefits, as we have seen.
But same as variables and indicators, reference levels also have a spatial resolution, and they should only be used to normalise variables when the variable is at a scale which is relevant to the way the reference levels are defined.
This was exeplified with the billberry example (@sec-bill).

@@

As we saw in the wolves example (@sec-wolves) this eases interpretation and reduces the change that the meaning of indicators are misunderstood.
It may be premature, or not even possible, to prescribe an aggregation pathway to be used for all ECA, but internal consistency within assessment should be possible.

>>>>

## Alien species example {#sec-ali}

@@

## Phytoplankton trophic index (PTI) {#sec-pti}

@@

### 1. Report the aggregation pathway, using standardised terminology {.unnumbered}

@@

### 2. Normalise variables early, but at the scale where the reference levels are relevant {.unnumbered}

Normalising variables early in the aggregation pathway has several benefits, such as comparability between nested assessments (@sec-ali).
It also means that you are aggregating normative measures of condition, which is generally what you want in ECAs.
But same as variables and indicators, reference levels also have a spatial resolution, and they should only be used to normalise variables when the variable is at a scale which is relevant to the way the reference levels are defined.
This was exemplified with the billberry example (@sec-bill).

@@

It may be premature, or not even possible, to prescribe an aggregation pathway to be used for all ECAs, but internal consistency within assessments should be possible.
As we saw in the wolves example (@sec-wolves), this eases interpretation and reduces the change that the meaning of indicators are misunderstood.
